Updated Criteria for the Diagnostic Procedure for Parkinson’s Disease Dementia on Level I and their Validity in Deep Brain Stimulation Cohort

### 2024-06-12 theart.qmd
Uploaded article, introduction added

---
title: "Updated Criteria for the Diagnostic Procedure for Parkinson's Disease Dementia on Level I"
shorttitle: "Parkinson's Disease Dementia Level I Criteria"
author:
  - name: Martina Kvapilova
    corresponding: false
    role:
      - Conceptualization
      - Data curation
      - Writing - original draft
    affiliations:
      - id: "1FM"
        name: "First Faculty of Medicine and General University Hospital in Prague, Charles University, Czech Republic"
        department: "Department of Neurology and Centre of Clinical Neuroscience"
  - name: Josef Mana
    corresponding: false
    orcid: "0000-0002-7817-3978" 
    email: "josef.mana@protonmail.com"
    role:
      - Conceptualization
      - Data curation
      - Investigation
      - Formal analysis
      - Software
      - Methodology
      - Project administration
      - Validation
      - Writing - original draft
    affiliations:
      - id: "1FM"
        name: "First Faculty of Medicine and General University Hospital in Prague, Charles University, Czech Republic"
        department: "Department of Neurology and Centre of Clinical Neuroscience"
  - name: Martin Cihak
    corresponding: false
    orcid: "0000-0001-7765-4799"
    email: "martin.cih@gmail.com"
    role:
      - Conceptualization
      - Investigation
      - Methodology
      - Writing - original draft
    affiliations:
      - id: "1FM"
        name: "First Faculty of Medicine and General University Hospital in Prague, Charles University, Czech Republic"
        department: "Department of Neurology and Centre of Clinical Neuroscience"
  - name: Ondrej Bezdicek
    corresponding: false
    orcid: "0000-0002-5108-0181"
    email: "ondrej.bezdicek@gmail.com"
    role:
      - Investigation
      - Data curation
      - Funding acquisition
      - Conceptualization
      - Project administration
      - Supervision
      - Writing - original draft
    affiliations:
      - id: "1FM"
        name: "First Faculty of Medicine and General University Hospital in Prague, Charles University, Czech Republic"
        department: "Department of Neurology and Centre of Clinical Neuroscience"
  - name: Tereza Uhrova
    corresponding: false
    email: "tereza.uhrova@vfn.cz"
    role:
      - Investigation
    affiliations:
      - id: "1FM"
        name: "First Faculty of Medicine and General University Hospital in Prague, Charles University, Czech Republic"
        department: "Department of Neurology and Centre of Clinical Neuroscience"
  - name: Robert Jech
    corresponding: false
    orcid: "0000-0002-9732-8947"
    email: "jech@cesnet.cz"
    role:
      - Funding acquisition
      - Resources
      - Writing - review & editing
    affiliations:
      - id: "1FM"
        name: "First Faculty of Medicine and General University Hospital in Prague, Charles University, Czech Republic"
        department: "Department of Neurology and Centre of Clinical Neuroscience"
abstract: ""
keywords:
  - Parkinson’s disease
  - dementia
  - level I criteria
format:
 apaquarto-pdf:
   documentmode: man
   a4paper: true
   pdf-engine: lualatex
   floatsintext: false
   keep-tex: false
bibliography: references.bib
---

@@

@@

To measure independence in activities of daily living, we administered Functional Activities Questionnaire (FAQ) [@bezdicek2011; @pfeffer1982]. To assess neuropsychiatric functioning, we used Beck Depression Inventory II (BDI) [@ciharova2020; @beck1996] and State-Trait Anxiety Inventory (STAI) [@spielberg1983; @mullner1980]. Psychotic symptoms were assessed in an interview by a trained psychiatrist.

>>>>

---
lang: en
title: "Updated Criteria for the Diagnostic Procedure for Parkinson’s Disease Dementia on Level I and their Validity in Deep Brain Stimulation Cohort"
shorttitle: "Parkinson's Disease Dementia Level I Criteria"
author:
  - name: Martina Kvapilova
    corresponding: false
    orcid: "0009-0007-4665-3946"
    role:
      - Conceptualization
      - Data curation
      - Writing - original draft
    affiliations:
      - id: "1FM"
        name: "First Faculty of Medicine and General University Hospital in Prague, Charles University, Czech Republic"
        department: "Department of Neurology and Centre of Clinical Neuroscience"
  - name: Josef Mana
    corresponding: false
    orcid: "0000-0002-7817-3978" 
    email: "josef.mana@protonmail.com"
    role:
      - Conceptualization
      - Data curation
      - Investigation
      - Formal analysis
      - Software
      - Methodology
      - Project administration
      - Validation
      - Writing - original draft
    affiliations:
      - id: "1FM"
        name: "First Faculty of Medicine and General University Hospital in Prague, Charles University, Czech Republic"
        department: "Department of Neurology and Centre of Clinical Neuroscience"
  - name: Martin Cihak
    corresponding: false
    orcid: "0000-0001-7765-4799"
    email: "martin.cih@gmail.com"
    role:
      - Conceptualization
      - Investigation
      - Methodology
      - Writing - original draft
    affiliations:
      - id: "1FM"
        name: "First Faculty of Medicine and General University Hospital in Prague, Charles University, Czech Republic"
        department: "Department of Neurology and Centre of Clinical Neuroscience"
  - name: Tereza Uhrova
    corresponding: false
    email: "tereza.uhrova@vfn.cz"
    role:
      - Investigation
    affiliations:
      - id: "1FM"
        name: "First Faculty of Medicine and General University Hospital in Prague, Charles University, Czech Republic"
        department: "Department of Neurology and Centre of Clinical Neuroscience"
  - name: Robert Jech
    corresponding: false
    orcid: "0000-0002-9732-8947"
    email: "jech@cesnet.cz"
    role:
      - Funding acquisition
      - Resources
      - Writing - review & editing
    affiliations:
      - id: "1FM"
        name: "First Faculty of Medicine and General University Hospital in Prague, Charles University, Czech Republic"
        department: "Department of Neurology and Centre of Clinical Neuroscience"
  - name: Ondrej Bezdicek
    corresponding: true
    orcid: "0000-0002-5108-0181"
    email: "ondrej.bezdicek@gmail.com"
    role:
      - Investigation
      - Data curation
      - Funding acquisition
      - Conceptualization
      - Project administration
      - Supervision
      - Writing - original draft
    affiliations:
      - id: "1FM"
        name: "First Faculty of Medicine and General University Hospital in Prague, Charles University, Czech Republic"
        department: "Department of Neurology and Centre of Clinical Neuroscience"
abstract: ""
keywords:
  - Parkinson’s disease
  - Parkinson's disease dementia
  - level I criteria
format:
 apaquarto-docx:
   documentmode: man
   a4paper: true
   pdf-engine: lualatex
   floatsintext: false
   keep-tex: false
bibliography: references.bib
---

@@

Parkinson’s disease (PD) is a neurodegenerative disorder characterised by the slow, progressive onset of rigidity, bradykinesia, and resting tremors that ultimately advance to motor and non-motor impairment [@postuma2015].

The diagnostic criteria for Parkinson’s disease dementia (PDD) were developed in 2006 and their procedure followed shortly after [@emre2007; @dubois2007]. At that period the PDD criteria were oriented at criteria originating in Alzheimer’s disease (AD) since there was a lack of PD biomarkers to promote a biological classification of PD [@yamashita2023; @höglinger2024].

The newer Parkinson’s disease mild cognitive criteria (PD-MCI) led to an advance in different aspects of PD research [@litvan2012]. These included progress in clinical, genetic, therapeutic modalities, clinical trials or conversion rates to PDD [@hoogland2017; @hoogland2019; @aarsland2021].

At present, there is an ongoing effort to revisit the PDD criteria to secure the comparability of studies and clinical trials across different sites [@kulisevsky2024]. In our previous research, we aimed at establishing rigorous psychometric procedure to differentiate PD-MCI from PDD [@bezdicek2017; @bezdicek2016]. Hence the diagnosis of PDD is an exclusion criterion in deep brain stimulation (DBS), we decided to test the revisited PDD criteria in our PD cohort selected for undergoing DBS [@deuschl2006].

The current study aims at showing the agreement and differences between establishing the diagnosis using only Level 1 criteria as suggested by previous diagnostic procedure (the Movement Disorders Society (MDS)-Task Force screening checklist for detecting PD) [@emre2007; @dubois2007] in comparison to newly revised [@kulisevsky2024] one to validate them. The second objective of our study was to determine the agreement of revised PDD criteria with full neuropsychological testing at Level 2.

@@

To measure independence in activities of daily living, we administered the Functional Activities Questionnaire (FAQ) [@bezdicek2011; @pfeffer1982]. To assess neuropsychiatric functioning, we used Beck Depression Inventory-II (BDI) [@ciharova2020; @beck1996] and State-Trait Anxiety Inventory (STAI) [@spielberg1983; @mullner1980]. Psychotic symptoms were assessed in an interview by a trained psychiatrist.
